---
title: "`r params$title`"
author: "Ryan Heslin"
date: "`r format(Sys.Date(), '%B %e, %Y')`"
header-includes:
  - \setlength{\parindent}{2em}
  - \setlength{\parskip}{2em}
params:
    title: "Chocolate Exploratory Data Analysis"
output:
  pdf_document:
    highlight: "kate"
    df_print: "kable"
    includes:
      in_header: "notes_text_preamble.tex"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "",
  fig.pos = "",
  message = FALSE,
  tidy = "styler",
  warning = FALSE,
  fig.align = "center",
  highlight = TRUE
)
```


The data for this project come from week 3 of 
2022 Tidy Tuesday, available [here](https://github.com/rfordatascience/tidytuesday/blob/master/data/2022/2022-01-18/readme.md). They consist of ratings for 
`nrow(chocolate)` chocolate bars. Useful features include country of origin, manufacturing company, 
ingredients, and words used by taste-testers to describe each bar. (How do I get that job?) I have preprocessed the data by extracting these descriptive words into columns and creating a dummy column for each ingredient.

```{r}
library(tidyverse)
source(here::here("R", "utils.R"))

chocolate <- readRDS(here::here("data", "chocolate.Rds"))
```

## The Response 


Chocolate bar quality is rated on a scale from 1 to 4 
inclusive, with increments of 0.25. The mean is 
`r mean(chocolate[["rating"]])`. I will have to 
round predictions to that scale, and I may have to 
create a custom loss function to accommodate it.
```{r}
vec2DF(table(chocolate[["rating"]]))
```

The cumulative distribution
```{r}
ggplot(chocolate, aes(x = rating)) +
  stat_ecdf(geom = "step", fill = "blue", alpha = .5) +
  labs(y = "Cumulative Proportion")
```

```{r}

```
The vast majority of companies are American, which 
may reflect bias in the data. 

The cacao beans used to make the chocolate bars come from just six countries.
```{r}
table_head(chocolate[["company_location"]])

table_head(chocolate[["country_of_bean_origin"]]) |>
  vec2DF()
```

# Flavor Description

I created five dummy columns to hold each word or phrase used to describe each bar. I drop the fifth because only one bar had that many terms

Most of the words seem to describe bar ingredients 
("oily", "sweet", etc.); fewer convey direct sentiments. They may still be useful as predictors 
if the testers consistently preferred some 
combinations of ingredients to others.

The worst-rated bars have some amusing words.
```{r}
chocolate <- select(chocolate, -characteristic5)

filter(chocolate, rating == 4) |>
  select(starts_with("characteristic")) |>
  head()

filter(chocolate, rating == 1) |>
  select(starts_with("characteristic"))
```

```{r}
all_words <- select(chocolate, starts_with("characteristic")) |>
  unlist() |>
  na.omit()

length(all_words)
library(tidytext)
get_sentiments("bing")

sum(all_words %in% (get_sentiments("bing")$word))
```


Of the `r all_words` total words recorded, only 
a fraction appear in a sentiment lexicon, though 
that's not too bad. Many are repeated, with only 
`r n_distinct(all_words)` distinct values.


The most common are generic flavor descriptors. 

```{r}
table_head(all_words, .n = 10) |>
  vec2DF()
ingredient_words <- names(table_head(all_words, .n = 10))
```


As we'd expect, almost all chocolate bars have 
cacao and sugar. The other ingredients are much 
rarer.
```{r}
chocolate |>
  select(beans:last_col()) |>
  colMeans() |>
  vec2DF() |>
  pivot_longer(everything(),
    names_to = "Ingredient",
    values_to = "Proportion With"
  ) |>
  mutate(Ingredient = factor(Ingredient, levels = unique(Ingredient)[order(`Proportion With`)])) |>
  ggplot(aes(y = Ingredient, x = `Proportion With`, fill = Ingredient)) +
  geom_col() +
  scale_x_continuous(labels = scales::percent) +
  labs(title = "Proportion of Chocolate Bars with Recorded Ingredients") +
  theme(legend.position = "none")
```

# Conclusion 

`country_of_bean_origin` and `company_location` would be useful features as is. It will take more work to wring usable data out of ingredients and text. Text could be transformed into sentiments, but that would not be useful for words describing 
flavors or ingredients.
